### 2015-08-24 presentations/paper/abstract.tex
Move sections into separate files

\begin{abstract}
  In 2014, the Git surpassed Subversion as the top code management tool
  \cite{EComSurv14}. With so many authors colaborating comes the need for a
  appropriate and standardized use of elemental actions like doing commits.
  Although the official Git documentation states very clear guidelines
  \cite{OffGuide} on how to do a commit, a lot of users are not aware of them or
  simply ignore them.

  This paper presents a tool that analyzes a set of commits regarding criteria
  composed of the official guidelines and some common mistakes, and shows a
  prototypical website that encourages improvements in commit-etiquette by
  creating a competitive score-board for a given GitHub repository. We also
  perform a machine learning powered analysis of correlations between all
  criteria.
\end{abstract}