### 2015-08-24 presentations/paper/details.tex

\section{Details}
\label{S:details}

Any sort of interesting details such as source code, extra data (extra
to the key results), extra procedural details supporting
reproducibility can be listed here.