### 2015-08-24 presentations/paper/introduction.tex

\section{Introduction}

Address these questions:

\begin{itemize}

\item What is the context of this research?

\item What is the overall challenge addressed by this research?

\item What is the overall approach adopted by the research?

\item Summarize results very briefly.

\end{itemize}