### 2015-08-24 presentations/paper/research_question.tex

\section{Research question}

State 1-5 research questions. One question could be enough. More is
not better.

For each question, make clear why this is an interesting question. For
example, try to make clear why a scientist or a practician would
benefit from knowing the answer to the question.

Lay out some background for your research questions. That is, what
body of related work may be worth mentioning and comparing to the
research at hand. This may be related in terms of both relevant
methodology or similar research questions.

You should make sure to carry out a proper related work study. Use
DBLP on the MSR conference for 2+
years.\footnote{\url{http://dblp.uni-trier.de/db/conf/msr/}} You are
welcome to search other conferences and journals. Coming to the
conclusion that there is no related work is absolutely
unacceptable. Again, you need to show related work awareness so that
your research question is substantiated as being relevant and
your methodology is deeply informed by existing publications.

In many cases, you want to refine your research questions into actual
hypotheses, i.e., something that is potentially falsifiable, subject
to an appropriate approach towards measuring.

It is also common to refine research questions. That is, one starts
with a general (overall) research question and takes it appart into
several questions that are concrete enough to be researched on and to
be answered. In this manner, you may often suffice with one or two
research questions which are however refined into perhaps
several questions. Again, even without refinement, just one question may be
Ok. This depends on the nature of the project and on the kind of question.

Here are some recommended MSR papers by the Software Languages
Team~\cite{LaemmelLPV11,LaemmelP13,SchmorleizL15} which you could
consult to see instances of the concepts discussed in this
metamodel. These papers may or may not be in an optimal form to
function as good advice. You are also advised to have a look at some
MSR papers (from the MSR conference), but again, they may also
slightly vary in terms of consistency of style. if you like to get
general advice on scientific writing, then I recommend relevant
textbooks~\cite{Day98,Zobel09}. The best way to learn writing is by
writing and by having someone with experience to provide feedback.

You should most definitely have a look at this excellent tutorial on
empirical software engineering (ESE)~\cite{Easterbrook07} to which MSR
relates. General ESE is quite often concerned with controlled
experiments and other research methods that involve subjects, whereas
MSR is more concerned with automated analysis, but the text is still
very helpful in coming to grips with research methods in software
engineering overall. Specifically, it helps with formulating research questions.