### 2015-08-24 presentations/paper/paper.tex
Use our title and authors

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{url}
\usepackage{comment}

\pagenumbering{arabic}

\begin{document}
\title{Commit Rater}
\subtitle{Analyzing and Comparing Git Author's Commit Messages}
\author{Carsten Hartenfels}% \and Marco Brack} TODO: This causes some error
\institute{Software Languages Team, CS Faculty, University of Koblenz-Landau}

\maketitle

\include{abstract}
\include{introduction}
\include{research_question}
\include{methodology}
\include{results}
\include{threads}
\include{conclusion}

\appendix
\include{details}

\bibliography{paper}
\bibliographystyle{abbrv}

\end{document}

\include{notes}