### 2015-08-25 presentations/paper/paper.tex
Finish data extraction aside from the time table

\documentclass{llncs}

\usepackage{hyperref}
\usepackage{url}
\usepackage{comment}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\renewcommand{\arraystretch}{1.5}

\pagenumbering{arabic}

\begin{document}
\title{Commit Rater}
\subtitle{Analyzing and Comparing Git Author's Commit Messages}
\author{Carsten Hartenfels \and Marco Brack}
\institute{Software Languages Team, CS Faculty, University of Koblenz-Landau}

\maketitle

\include{abstract}
\include{introduction}
\include{research_question}
\include{methodology}
\include{results}
\include{threads}
\include{conclusion}

\appendix
\include{details}

\bibliography{paper}
\bibliographystyle{abbrv}

\end{document}

\include{notes}