### 2015-08-25 presentations/paper/introduction.tex
Comment out introduction info

\section{Introduction}

%Address these questions:
%
%\begin{itemize}
%
%\item What is the context of this research?
%
%\item What is the overall challenge addressed by this research?
%
%\item What is the overall approach adopted by the research?
%
%\item Summarize results very briefly.
%
%\end{itemize}