### 2015-08-25 presentations/paper/threats.tex
Renamed threads to threats

\section{Threats to validity}
\label{sec:threats-to-validity}

There are external and internal threats to validity. See the softlang
papers and MSR papers for examples. See this resource~\cite{Michael04}
for an explanation. Be concise and systematic about answering
questions.