### 2015-08-25 presentations/paper/conclusion.tex
Comment out instructions in Conclusion

\section{Conclusion}
\label{sec:conclusion}

%This is essentially a redundant section. Try to summarize your
%research from a more informed point of view. (We assume that the
%reader has read the paper and is ready for some deeper conclusion.)
%You could very well add some perspectives for future work, without
%repeating smaller scale weakness of research, as they were already
%identified in the threats to validity section.