### 2015-08-26 presentations/paper/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we presented an application used to rate commit messages of Git repositories. Data extraction is done by cloning the repositories. Data synthesis is based on several separate criteria -- some official, some set by us personally. Finally, data analysis is done by comparing the various authors and repositories against each other\TODO{is that true?}.

Both research questions were successfully answered: commits were recognized by the aformentioned criteria and by using machine learning, we successfully validated our personal criteria by correlating it with the criteria from reputable sources\TODO{maybe fix when it's done}.

A prototype web application has been developed and is available for public testing, allowing us to use the resulting data for further research and development.

\subsection{Future Work}
\label{sec:future-work}

%This is essentially a redundant section. Try to summarize your
%research from a more informed point of view. (We assume that the
%reader has read the paper and is ready for some deeper conclusion.)
%You could very well add some perspectives for future work, without
%repeating smaller scale weakness of research, as they were already
%identified in the threats to validity section.