### 2015-08-26 presentations/paper/details.tex
Write something in the details section

\section{Details}
\label{sec:details}

The source code for the Commit Rater can be found on GitHub\footnote{\url{https://github.com/hartenfels/Commit-Rater}}. The repository includes all of the boxplots (\texttt{analysis/boxplots}), their raw data source data (\texttt{samples}) and their generation script (\texttt{analysis/boxplots.rb).

\TODO{It also includes the files used for the machine learning...}

Finally, the web frontend runs on a private server\footnote{\url{http://trollarena.de:8080/}}, it may go offline in the not too distant future. Its source code is also available on GitHub\footnote{\url{https://github.com/hartenfels/Commit-Rater-Web}}.