### 2015-08-26 presentations/paper/abstract.tex

\begin{abstract}
  In 2014, Git surpassed Subversion as the top code management tool
  \cite{EComSurv14}. With many authors collaborating comes the need for an
  appropriate and standardized use of elemental actions, like formulating
  commit messages. Although the official Git documentation states very clear
  guidelines \cite{OffGuide} on how this should be done, many users are either
  unaware or simply ignore them.

  This paper presents a tool that analyzes a set of commit messages regarding
  criteria composed of the official guidelines and some common mistakes, and
  shows a prototypical website that encourages improvements in commit message
  style by creating a competitive score-board for a given GitHub repository. We
  also perform a machine-learning-powered analysis of correlations between the
  criteria, which can help improve the selection of criteria in the future.
\end{abstract}