### 2015-09-25 presentations/paper/details.tex
fix bug

\section{Details}
\label{sec:details}

The source code for the Commit Rater can be found on GitHub\footnote{\url{https://github.com/hartenfels/Commit-Rater}}. The repository includes all of the boxplots (\texttt{analysis/boxplots}), their raw data source data (\texttt{samples}) and their generation script (\texttt{analysis/boxplots.rb}).

Weka's machine learning files (ARFF) can also be found in the (\texttt{samples}) folder, including a combined source file of all sample repos and the output of the Weka associator for it.

Finally, the web frontend runs on a private server\footnote{\url{http://trollarena.de:8080/}}. It may go offline in the not too distant future. Its source code is also available on GitHub\footnote{\url{https://github.com/hartenfels/Commit-Rater-Web}}.